Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-17 writeups/Su_disk_disp.tex
recolor plots for clarity

Since $\p*{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\autoref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\autoref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by Eqs.~\eqref{se:area_ward}. The dotted lines correspond to
    small $\eta$ approximations discussed in Appendix~\ref{s:ad_approx}. The
    dashed lines for $\eta > \eta_{\rm c}$ are the effective values of $A_{\rm
    II}, A_{\rm III}$ for $\eta > \eta_{\rm c}$, denoting the points that would
    flow into either area under adiabatic decrease of $\eta$ from $\eta >
    \eta_{\rm c}$. The vertical lines correspond to $\eta = \eta_{\rm c}$
    (Eq.\eqref{eq:etac}) and the values of $\eta$ for which $A_{\rm II}$ is
    maximized ($\eta_{\rm \max, II}$) and for which $A_{\rm III}$ is minimized
    ($\eta_{\rm \min, III}$, Eq.~\eqref{eq:eta_minIII}).}\label{fig:eq_areas}
\end{figure}

>>>>

Since $\p*{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\autoref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\autoref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{The solid lines show the fractional areas of each of the zones
    $A_{\rm i}(\eta) / 4\pi$ as given by Eqs.~\eqref{se:area_ward}. The colored
    dashed lines correspond to small $\eta$ approximations used in
    Appendix~\ref{s:ad_approx}. The colored dashed lines for $\eta > \eta_{\rm
    c}$ are the effective values of $A_{\rm II}, A_{\rm III}$ for $\eta >
    \eta_{\rm c}$, denoting the points that would flow into either area under
    adiabatic decrease of $\eta$ from $\eta > \eta_{\rm c}$. The vertical black
    dashed lines correspond to $\eta = \eta_{\rm c}$ (Eq.\eqref{eq:etac}) and
    the values of $\eta$ for which $A_{\rm II}$ is maximized ($\eta_{\rm \max,
    II}$) and for which $A_{\rm III}$ is minimized ($\eta_{\rm \min, III}$,
    Eq.~\eqref{eq:eta_minIII}). The horizontal black dashed line illustrates the
    fraction of phase space occupied by zone II at separatrix appearance.
    }\label{fig:eq_areas}
\end{figure}
